Linear and Differential Cryptanalysis of DES

### 2009-01-30 trunk/talk/draft1.tex
More Linear SPN slides

>>>>

\begin{frame}
Now, come back to Per-round approximations. We can write down 4 approximations (Let $S_{i,j}$ repesent the $j$-th S-Box in the $i$-th round):
\begin{eqnarray*}
S_{1,2}: X_1 \oplus X_3 \oplus X_4 & = & Y_2 \\
S_{2,2}: X_2 = Y_2 \oplus Y_4 \\
S_{3,2}: X_2 = Y_2 \oplus Y_4 \\
S_{3,4}: X_2 = Y_2 \oplus Y_4 \\
\end{eqnarray*}
Each of these has a probability bias magnitude of $1/4$. We can use the Piling Up Principle to combine them into a single expression relating plaintext bits to cyphertext bits.
\end{frame}

\begin{frame}
Consider the first 2 rounds:\\
\vspace{5mm}
\[  X_1 \oplus X_3 \oplus X_4 = Y_2 \Rightarrow \]
\[ V_{1,6} = (P_5\oplus K_{1,5}) \oplus (P_7\oplus K_{1,7}) \oplus (P_8\oplus K_{1,8}) \]
\vspace{2mm}
\[ X_2 = Y_2 \oplus Y_4 \Rightarrow \]
\[ V_{2,6} \oplus V_{2,8} = V_{1,6} \oplus K_{2,6} \]
\vspace{2mm}
Each of these has a bias of magnitude $1/4$. We can combine:
\[  V_{2,6} \oplus V_{2,8} \oplus P_5 \oplus P_7 \oplus P_8 \oplus K_{1,5} \oplus K_{1,7} \oplus K_{1,8} \oplus K_{2,6} = 0 \]
By the Piling Up Principle this expression holds with bias:
\[ 2 * 1/4 * 1/4 = 1/8 \]
\end{frame}

\begin{frame}
Using this principle we can write the following equation over 3 rounds of SPN:
\[ U_{4,6} \oplus U_{4,8} \oplus U_{4,14} \oplus U_{4, 16} \oplus P_5 \oplus P_7 \oplus P_8 = \Sigma K \]
Where $\Sigma K$ is the sum over some key bits. Note that Since the key is fixed, $\Sigma K = 0$ or $1$ and thus we can ignore it since we only care about the bias.\\
\vspace{5mm}
The magnitude of the bias of the above expression, by the Piling Up Principle, is $1/32$. \\
\vspace{5mm}
Next we show how we can extract key bits using this information.
\end{frame}

\begin{frame}
Attack Idea:\\
\vspace{3mm}
We have an expression that links plaintext bits to input bits to the 4th round of SPN that holds with high bias ($1/32$). We can partially undo the last round by guessing the last key. \\
\vspace{3mm}
\pause
If we guess correctly, the equation will hold with high bias. If we guess wrong, the equation will probabily hold with probability close to 1/2, that is, a bias close to 0.\\
\vspace{3mm}
BUT To do this, we don't need to guess the entire key for the last round! Our expression only involves 4 key bits, output from 2 S-Boxes. Thus we only need to guess $2^8 = 256$ values.\\
\vspace{3mm}
For each value of the guessed \textit{target partial subkey} we can undo the last round and determine the bias of the equation. Highest bias indicates likely correct guess.
\end{frame}
